---
title: "analysis"
---

##Starting out
###Quarto exercices

Loading the libraries.

```{r}
#| output: false

library(babynames)
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(pheatmap)
```

Showing the head of the `babynames` dataset

```{r}
head(babynames) |> kable()
```

Two functions to get the most frequent babynames over a time-period and to plot the top n most popular names.

```{r}
#| code-fold: true

get_most_frequent <- function(babynames, select_sex, from = 1950) {
  most_freq <- babynames |>
    filter(sex == select_sex, year > from) |>
    group_by(name) |>
    summarise(average = mean(prop)) |>
    arrange(desc(average))
    
  return(list(
    babynames = babynames,
    most_frequent = most_freq,
    sex = select_sex,
    from = from))
}

plot_top <- function(x, top = 10) {
  topx <- x$most_frequent$name[1:top]
  
  p <- x$babynames |>
    filter(name %in% topx, sex == x$sex, year > x$from) |>
    ggplot(aes(x = year, y = prop, color = name)) +
    geom_line() +
    scale_color_brewer(palette = "Paired") +
    theme_classic()
  
  return(p)
}
```

Plotting the top n names for girls:

```{r}
#| label: fig-top_girl_names
#| fig-cap: "Top girl names "

get_most_frequent(babynames, select_sex = "F") |> 
  plot_top()
```

Wow, @fig-top_girl_names looks amazing.

Plotting the top n names for boys:

```{r}
#| label: fig-top_boy_names
#| fig-cap: "Top boy names "

get_most_frequent(babynames, select_sex = "M") |> 
  plot_top()
```

Insane how interesting @fig-top_boy_names is.

###GitHub exercise section

```{r}
#| label: fig-git_plots
#| fig-cap: "Two line plots and a heatmap"
#| fig-subcap:
#|  - "Plot 1"
#|  - "Plot 2" 
#|  - "Heatmap"
#| layout: [[45, -10, 45], [100]]


# get most frequent girl names from 2010 onwards
from_year <- 2010
most_freq_girls <- get_most_frequent(babynames, select_sex = "F",
                                     from = from_year)

# plot top 5 girl names
most_freq_girls |>
  plot_top(top = 5)

# plot top 10 girl names
most_freq_girls |>
  plot_top(top = 10)

# get top 30 girl names in a matrix
# with names in rows and years in columns
prop_df <- babynames |> 
  filter(name %in% most_freq_girls$most_frequent$name[1:30] & sex == "F") |>
  filter(year >= from_year) |> 
  select(year, name, prop) |>
  pivot_wider(names_from = year,
              values_from = prop)

prop_mat <- as.matrix(prop_df[, 2:ncol(prop_df)])
rownames(prop_mat) <- prop_df$name

# create heatmap
pheatmap(prop_mat, cluster_cols = FALSE, scale = "row")
```
In @fig-git_plots-1 and @fig-git_plots-2 the line plots are shown. To view trends of many names, @fig-git_plots-3 displays a heatmap.